Eagle: Making multiple-locus association mapping on a genome-wide scale routine

### 2018-10-04 paper.tex
revisions

\documentclass{article}
\usepackage{xr}
\bibliographystyle{naturemag}
\renewcommand{\baselinestretch}{1.3} 
\usepackage{graphicx}

@@

%%      To Do
%%    1. is tense correct in methods - multiple and single locus mehtods
%%    2. add bold font to matrices and vectors
%%    3. change plot titles to be more condense
%%    4. fix box bounding problem with figure 1.

@@

\subsection{Simulation Study}
A large simulation study was performed where we sought to  answer two questions. 
First, how does Eagle compare, in terms of run time and memory usage, to 
competing implementations? Second, how well does Eagle find true associations (power) and avoid 
false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 X 400K),  1,500 individuals and 
50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).   
These scenarios reflect, at least in some cases, the sizes of study being performed in animals, plants, and humans.

@@

Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM. Not all data generated under the six scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs/packages was the limiting factor (See {\bf Supplementary Figure \ref{supfigmem}}).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
Of the multiple-locus programs/packages, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from our largest scenario, 10000 x 1.5M.

@@

\label{fig_time}
\begin{center}
\includegraphics[width=14cm, height=12cm]{Figure2_time.jpg}
\end{center}

@@

Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
have been performed and say the 500023rd, 15th, and 420th
SNP were selected. Then $S=\{500023, 15, 420\}$. 
Let $y^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
Let $M^{(n_g \; \times \; L)} = [m_1 m_2 \ldots m_L]$ be a matrix containing the genotype data which have been collected 
from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
may be taken of the same group/line/strain. 
 It is common for the columns of $M$ to be in map order but this is not a requirement. 
The vector $m_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.

The specifics of the Eagle method are as follows. 
Eagle builds the "best" model iteratively, via forward selection. 
Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
To perform the $s+1$th  iteration, we first fit the current model to the data. 
The (current) model is of the form 
\begin{equation}
\label{eq1}
y = X \tau + Z u_g + e
\end{equation}
where 
$X^{(n \; \times \; p)}$ and $Z^{( n \; \times \; n_g)}$ are known design matrices with $X$ being of full rank and $Z$ 
containing zeros and ones that assign the appropriate genetic effect to each measurement. 
The vector 
$\tau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
$u_g^{(n_g \; \times \; 1)}$ contains the 
genetic effects. The vector of residuals is 
$e^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(0, \sigma^2_e I^{(n \; \times \; n)})$. 
So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
However, 
it is how we specify $u_g$ that distinguishes our model from the others.

The genetic effects $u_g$ are modelled as 
\begin{equation}
\label{eq2}
u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + M_{-S} a_{-S}
\end{equation}
where $m_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
$a_{S_k}$ is the additive effect of the $k$th selected SNP, $M_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
with the data for the SNP in $S$ removed,  and $a_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
$a_{-S} \sim N(0, \sigma_a^2 I^{(L-s \; \times \;  L-s)})$. 
The terms in the summation on the left hand side are fixed effects.   They account 
for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
This is a simple genetic model but it 
is effective for discovering SNP-trait associations.

Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
can be computationally demanding. However, our model only contains a single random effect ($a_{-S}$). Here, highly efficient single-dimension 
optimisation via spectral decomposition is possible \cite{kang2008efficient}.

@@

\subsubsection{Reducing the dimension of the model}
In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
The vector $a_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
\begin{equation}
\label{eq3}
u_g = \sum_{k=1}^s  m_{S_k} a_{S_k} + (M_{-S} M_{-S}^T)^{1/2} a^*_{-S}
\end{equation}
where $a^* \sim N(0, \sigma_a^2 I^{(n_g \; \times \;  n_g)})$, and 
$(M_{-S} M_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
Although it may not be obvious, the two models are equivalent, 
having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
effects needing estimating.

Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\widetilde{a}$ from estimates from model  (\ref{eq3})  with 
\begin{equation}
\widetilde{a} = \left [ M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \right ] \widetilde{a}^*
\end{equation}
where its variance matrix is
\begin{equation}
\label{eq4}
\textrm{var}(\widetilde{a}) = M_{-S}^T (M_{-S} M_{-S}^T)^{-1/2} \textrm{var}(\widetilde{a}^*) (M_{-S} M_{-S}^T)^{-1/2} M_{-S}
\end{equation}
These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
of (\ref{eq4}).

@@

The data are generated via data perturbation (Zhao et al. 2007). Data perturbation amalgamates real with simulated data to generate replicates. 
It is a way of introducing greater realism into a simulation study. 
 Here, the genotype data are real, the 
quantitative trait data are simulated. 
 The SNP genotypes are drawn, 
according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.

To generate the trait data $y$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
generate a $g^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
Fifth, $e^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $y =  g + e$.  
In forming $y$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.

@@

Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
via C++ functions that utilise Eigen C++ routines (ref). Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
a link to our server for demonstrating Eagle on some test data are available on the Eagle website (http://eagle.r-forge.r-project.org).  
Eagle is available for download from the CRAN website (https://cran.r-project.org).

@@

Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.  Permutation and its variants \cite{ browning2008presto, pahl2010permory  } are also equally valid empirical approaches. Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the 
data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
 Instead,  analyses are 
performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
However, we would like to think that the parameter settings we used are sensible 
since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.

Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
including the extra random effects, is 
captured in this first-stage model. In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis. Even though this is a well accepted practice, it is approximate (ref).  A better solution is to fit a single model to the data. 
Although not specifically designed for association mapping,
WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RRGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
models,  run time and memory usage can become limiting factors for analysis.

>>>>

\documentclass{article}
\usepackage{xr}
\bibliographystyle{naturemag}
\renewcommand{\baselinestretch}{1.3} 
\usepackage{graphicx}
\usepackage{bm}

@@

%%      To Do
%%   1. make ARi's changes to paper
%%    2. add bold font to matrices and vectors
%%    3. change plot titles to be more condense
%%    4.tidy up doc - reformat - remove comments make nice
%%    5 write cover letter
%%    6 send cover letter to dan
%%    7 send paper to Josh
%%    8 find reviewer

\newcommand{\bwidetildea}{\widetilde{\bm{a}}}

\newcommand{\bu}{\bm{u}}
\newcommand{\be}{\bm{e}}
\newcommand{\btau}{\bm{\tau}}
\newcommand{\ba}{\bm{a}}
\newcommand{\bzero}{\bm{0}}
\newcommand{\bI}{\bm{I}}

\newcommand{\bX}{\bm{X}}
\newcommand{\by}{\bm{y}}
\newcommand{\bZ}{\bm{Z}}
\newcommand{\bM}{\bm{M}}
\newcommand{\bmm}{\bm{m}}

@@

\subsection{Simulation Study}
A large simulation study was performed where we sought to  answer two questions. 
First, how does Eagle compare, in terms of run time and memory usage, to 
competing implementations? Second, how well does Eagle find true associations (power) and avoid 
false associations (type 1 errors)? Data were generated under six different scenarios; a study of size 150 individuals 
and 5,000 single nucleotide polymorphisms (SNPs) (150 x 5K),  350 individuals and 400,000 SNPs (350 x 400K),  1,500 individuals and 
50,000 SNPs (1500 x 50K), 2,000 individuals and 500,000 SNPs (2000 x 500K), 4,000 individuals and 
1,500,000 SNPs (4000 x 1.5M), and 10,000 individuals and 1,500,000 SNPs (10000 x 1.5M).   
These scenarios reflect, at least in some cases, the sizes of study being performed in animals, plants, and humans.

@@

Memory usage and run (or elapse) times were recorded for Eagle and the other computer programs/packages across the simulation scenarios. 
Analyses were performed on a high-end desktop computer with dual 8-core Xeon processors and 128 gigabytes of RAM. Not all data generated under the six scenarios could be analysed by all implementations. Memory usage 
for many of the computer programs/packages was the limiting factor (See Supplementary Figure \ref{supfigmem}).  The single-locus program GEMMA was by 
far the most memory efficient. Not surprisingly, the multiple-locus programs were memory intensive. Most required in 
excess of the 128 gigabytes of available RAM for the analysis of data generated under 4000 x 1.5M and 10000 x 1.5M.  
Even FaST-LMM, when all the SNP data were being used to calculate the similarity matrix, ran out of memory for the larger scenarios.
Of the multiple-locus programs/packages, only Eagle,  
with its ability to handle data larger than the memory capacity of the computer, was capable of producing findings 
for data from our largest scenario, 10000 x 1.5M.

@@

\label{fig_time}
\centering
    \includegraphics[width=1.0\textwidth,natwidth=610,natheight=642]{Figure2_time.jpg}
\end{figure}

@@

Let $S = \{ S_1, S_2, \ldots, S_s\}$ be a set of $s$ ordinal numbers where $S_k$ is the $S_k$th ordered SNP that was 
selected in the $k$th iteration of the model building process. Suppose three iterations  $(s=3)$
have been performed and say the 500023rd, 15th, and 420th
SNP were selected. Then $S=\{500023, 15, 420\}$. 
Let $\by^{(n \; \times \;1)}$ be a vector containing $n$ measurements of the quantitative trait. 
Let $\bM^{(n_g \; \times \; L)} = [\bmm_1 \bmm_2 \ldots \bmm_L]$ be a matrix containing the genotype data which have been collected 
from $L$ loci that span the genome on $n_g$ groups/lines/strains.  Here, $n \geq n_g$ meaning that a single or several trait measurements 
may be taken of the same group/line/strain. 
 It is common for the columns of $\bM$ to be in map order but this is not a requirement. 
The vector $\bmm_j^{(n_g \; \times \; 1)}$ contains the genotypes for the $j$th SNP. 
The genotypes are coded as -1, 0, and 1 corresponding to SNP genotypes AA, AB, and BB, respectively.

The specifics of the Eagle method are as follows. 
Eagle builds the "best" model iteratively, via forward selection. 
Suppose $s$ iterations of our model building process have already been performed. This means $s$ SNP-trait 
associations have been identified.  It also means that $s$ separate genomic regions of interest have been found.  
To perform the $s+1$th  iteration, we first fit the current model to the data. 
The (current) model is of the form 
\begin{equation}
\label{eq1}
\by = \bX \btau + \bZ \bu_g + \be
\end{equation}
where 
$\bX^{(n \; \times \; p)}$ and $\bZ^{( n \; \times \; n_g)}$ are known design matrices with $\bX$ being of full rank and $\bZ$ 
containing zeros and ones that assign the appropriate genetic effect to each measurement. 
The vector 
$\btau^{(p \; \times \; 1)}$ has $p$ fixed effects parameters including the intercept. The vector 
$\bu_g^{(n_g \; \times \; 1)}$ contains the 
genetic effects. The vector of residuals is 
$\be^{(n \; \times \;1)}$ whose distribution is assumed to follow $N(\bzero, \sigma^2_e \bI^{(n \; \times \; n)})$. 
So far,  this model differs little from standard linear mixed models for association mapping \cite{yu2006unified,zhao2007arabidopsis} 
However, 
it is how we specify $\bu_g$ that distinguishes our model from the others.

The genetic effects $\bu_g$ are modelled as 
\begin{equation}
\label{eq2}
\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} +\bM_{-S} \ba_{-S}
\end{equation}
where $\bmm_{S_k}^{(n_g \; \times \; 1)}$ is the vector of genotypes for the  $k$th selected SNP, 
$a_{S_k}$ is the additive effect of the $k$th selected SNP, $\bM_{-S}^{(b \; \times \; L-s)}$ is the matrix of  SNP genotypes 
with the data for the SNP in $S$ removed,  and $\ba_{-S}^{(L-s \; \times  \; 1)}$ is a random effect whose distribution is 
$\ba_{-S} \sim N(\bzero, \sigma_a^2 \bI^{(L-s \; \times \;  L-s)})$. 
The terms in the summation on the left hand side are fixed effects.   They account 
for the additive effects of those SNPs that have been found to be in association with the trait. The other term is a random effect. 
It accounts for the joint effect of the yet-to-be-identified SNP that are in association with the trait. 
This is a simple genetic model but it 
is effective for discovering SNP-trait associations.

Second, we estimate the parameters of (\ref{eq1}) and (\ref{eq2}) via residual maximum likelihood (REML).  For complex models, REML
can be computationally demanding. However, our model only contains a single random effect ($\ba_{-S}$). Here, highly efficient single-dimension 
optimisation via spectral decomposition is possible \cite{kang2008efficient}.

@@

\subsubsection{Reducing the dimension of the model}
In practice, estimating the parameters of (\ref{eq2}) can be demanding, computationally. 
The vector $\ba_{-S}$ has $L-s$ random effects where in modern genome-wide association studies, 
$L$, the number of SNPs, can be extremely large.  An alternative model is given by 
Verbyla \cite{verbyla2012rwgaim,verbyla2014whole}. 
They show how to reformulate (\ref{eq2}) to be a model with a random effect with only $n$ elements
\begin{equation}
\label{eq3}
\bu_g = \sum_{k=1}^s  \bmm_{S_k} a_{S_k} + (\bM_{-S} \bM_{-S}^T)^{1/2} \ba^*_{-S}
\end{equation}
where $\ba^* \sim N(\bzero, \sigma_a^2 \bI^{(n_g \; \times \;  n_g)})$, and 
$(\bM_{-S} \bM_{-S}^T)^{1/2}$ can be calculated via singular value decomposition \cite{golub2012matrix}.  
Although it may not be obvious, the two models are equivalent, 
having identical variance structures. Yet, the computational cost of model (\ref{eq3}) compared to 
model (\ref{eq2}) is much less, due to the random term in model (\ref{eq3}) having only $n$ instead of $L-s$ 
effects needing estimating.

Verbyla \cite{verbyla2012rwgaim,verbyla2014whole} go on to show how to recover $\bwidetildea$ from estimates from model  (\ref{eq3})  with 
\begin{equation}
\bwidetildea = \left [ \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \right ] \bwidetildea^*
\end{equation}
where its variance matrix is
\begin{equation}
\label{eq4}
\textrm{var}(\bwidetildea) = \bM_{-S}^T (\bM_{-S} \bM_{-S}^T)^{-1/2} \textrm{var}(\bwidetildea^*) (\bM_{-S} \bM_{-S}^T)^{-1/2} \bM_{-S}
\end{equation}
These values are needed in order to calculate the score statistic $t_j^2$ for identifying the SNP in strongest association with the trait. 
Fortunately, when calculating $t_j^2$, only the diagonal elements of the variance matrix are needed which simplifies the  calculation 
of (\ref{eq4}).

@@

The data are generated via data perturbation \cite{zhao2007arabidopsis}. 
Data perturbation amalgamates real with simulated data to generate replicates. 
It is a way of introducing greater realism into a simulation study. 
 Here, the genotype data are real, the 
quantitative trait data are simulated. 
 The SNP genotypes are drawn, 
according to the specifications of the scenario, from data collected from the 1000 Genome Project, version 3   \cite{10002010map}. 
Across scenarios, the SNP data differs. Across replicates within a scenario, the SNP data are the same.

To generate the trait data $\by$, first, $q$, the number of SNPs that are to be assigned a quantitative value is drawn from a Poisson distribution with 
mean 30. Second, $q$ SNP are selected randomly. Third, we assume an additive model for the SNPs. The SNP genotypes AA, AB, and BB 
are assigned the values -1, 0, and 1, respectively. Fourth, the SNP effects are summed across the $q$ selected loci, for each individual, to 
generate a $\bm{g}^{(n \times 1)}$ vector of genetic values where $n$ is the number of individuals. 
Fifth, $\bm{e}^{(n \times 1)}$, a vector of residuals, is drawn from a normal distribution where $e_i \sim N(0, \sigma^2_e)$ and $\sigma^2_e$ is 
the residual variance that has been set to yield a trait with heritability 0.5. Sixth,  the trait data are formed as $\by =  \bm{g} + \bm{e}$.  
In forming $\by$, we have purposely not included any other environmental variables such as age, sex, or experimental design effects. This is because 
not all the methods were implemented to handle the inclusion of additional fixed effects. A two-stage modelling approach 
is often adopted to deal with this situation, but we chose not to introduce this complexity into the analyses.

@@

Eagle has been implemented as an R package of the same name. Much of the computation though is performed outside of R 
via C++ functions that utilise Eigen C++ routines. Eagle has been purpose built to rely heavily on calls to BLAS and LAPACK, 
mathematical libraries common to most computer systems. By making use of multi-threaded  BLAS and LAPACK libraries, many of the 
calculations in Eagle are parallelised. We have gone to great lengths to make Eagle easy-to-use. Tutorials, videos, How-To guides, and 
a link to our server for demonstrating Eagle on some test data are available on the Eagle website (http://eagle.r-forge.r-project.org).  
Eagle is available for download from the CRAN website (https://cran.r-project.org).

@@

Eagle outperformed the other multiple-locus methods in our simulation study. However, we are cognizant of the fact  that we made several implementation 
choices that impact our conclusions.  For instance, we chose to calculate the significance of the 
SNP effects from bigRR, LMM-Lasso, and glmnet via stability selection.  Permutation and its variants \cite{ browning2008presto,pahl2010permory} are also equally valid empirical approaches. Stability selection though has the advantage of being based on repeated sampling of only a proportion (50\% in our case) of the 
data. Also, when analysing the (sub)samples, it was not necessary to calculate the entire solution path for a method. 
 Instead,  analyses are 
performed for a fixed value of the regularisation parameter, greatly reducing the amount of computation required. For r2VIM, an R package 
implementing random forests, we had to decide on the  minimum size of a terminal node, the number of trees, and number of potential variables. 
The setting of these parameters greatly affects performance.  We  acknowledge that in the hands of an expert, 
r2VIM could be fine-tuned for a better balance of computational and statistical performance. 
However, we would like to think that the parameter settings we used are sensible 
since they match the values in the original r2VIM publication \cite{szymczak2016r2vim}.

Eagle's computational speed does come at a cost. It is a weakness shared by all of the methods considered here, although in different ways. 
Eagle cannot handle extra random effects which are sometimes needed when more advanced study designs are employed. One solution 
is to adopt a two-stage analysis procedure. In the first stage, a single linear mixed model is fitted to the data. Much of the modelling complexity, 
including the extra random effects, is 
captured in this first-stage model. In the second stage, Eagle is run not on the original trait data but adjusted trait data which are obtained from the first stage analysis. Even though this is a well accepted practice, it is approximate (ref).  A better solution is to fit a single model to the data. 
Although not specifically designed for association mapping,
WGAIM \cite{verbyla2007analysis}, upon which Eagle is based, and RWGAIM \cite{verbyla2012rwgaim}  are two R packages where this is possible. The difficulty is that for large datasets and/or complex 
models,  run time and memory usage can become limiting factors for analysis.
